%!TEX root = ../main.tex

\chapter{Introduction}
\label{chp:intro}

\ac{NTNs} denote a category of networks where at least one link is routed via an aerial or space-borne vehicle such as \ac{HAPs}, \ac{UAVs} or telecommunication satellites.

\section{The focus on non-terrestrial networks}
The Third Generation Partnership Project \footnote{\href{https://www.3gpp.org}{\texttt{3gpp.org}}} (3GPP), the standardization body developing protocols for mobile communication networks, recently put a great emphasis on the importance of the integration of different access technologies with the existing terrestrial mobile telecommunication infrastructure \cite{3gpp-tr-21.917}. The envisioned future for mobile communications, starting with the already established 5G \ac{NR} and expanding with the new sixth-generation cellular networks, foresees the integration of a non-terrestrial component. The latest releases (Rel. 17 and Rel. 18) require that 5G and 6G systems shall be able to provide connectivity exploiting non-terrestrial, satellite access \cite{overview-rel-17-18-saad} \cite{5g-nr-communication-geo-leo-maattanen}.

\subsubsection{Limitations of terrestrial networks}
\paragraph{Remote places}
While \ac{TN} make the well-established foundation of today’s mobile communication infrastructure, their own nature poses some intrinsic limitations to their deployment in certain scenarios, especially in rural and remote areas. Conditions such as harsh terrain and hard to reach places act as natural barriers to the deployment of terrestrial infrastructure, which also requires the presence of an already established reliable power grid, driving up the costs that telecommunication companies would have to sustain.
Furthermore, the population density is often low in remote and rural places, making this kind of market particularly unattractive to private investors, further limiting the possibility for the people living there to access a resource which is becoming increasingly more important.

As studied and documented in \cite{6g-challenge-opportunity-base-pyramid}, the issue of an inadequate broadband coverage to rural regions is an enormous challenge, but also a great opportunity to kick start the economy of currently underdeveloped countries, promoting a more fair access to the internet and alleviating the infamous digital divide between different parts of the world.
\paragraph{Redundancy}
Another limitation of the current terrestrial infrastructure is the lack of robustness against natural disasters. Extreme events such as earthquakes, fires and floods, but also deliberate behaviours such as targeted attacks by terrorist organizations and sabotages can disrupt the connectivity for a prolonged duration, hindering rescue efforts and causing significant economical damage.
In this scenario, \ac{NTNs} can act as redundant access methodology, both to reduce downtimes of terrestrial infrastructure and to provide additional capacity when required. 

\paragraph{Long distances}
Remote equipment, offshore plants and distribution grids will also benefit from the research carried out in this field, since providing terrestrial connectivity in those scenarios would be a challenging task. The installation of an optical fiber link to serve a single endpoint, such as a wind turbine powered power plant in the ocean, or where many sensors needing a connection are placed in a large area, would be excessively costly \cite{performance-ntn-support-iot-wang}.

Other scenarios where \ac{NTNs} can become useful in overcoming the limitation of a terrestrial network are well descripted in \cite{ntn-6g-era-challenges-giordani} and \cite{potential-multilayered-nierarchical-ntn-wang}.

\subsection{Satellite types}
Depending on the orbit, satellites are classified in three main categories: \ac{GEO}, \ac{MEO} and \ac{LEO}.
\begin{itemize}
    \item \textbf{\ac{GEO} satellites}: orbiting at 35.786Km, \ac{GEO} satellites appear stationary since their orbiting period is the same as the Earth rotation period. This simplifies the tracking, and each satellite can provide connectivity to a fixed area.
    Their higher altitude creates a large cell footprint, larger than both MEO and LEO, so the cost per coverage area is lower.

    Cons are mainly linked to the large distance between the \ac{UE} and the satellite: the transmission power and the antenna gain have to be higher to account for the greater propagation losses, and the propagation delay of the signal travelling at the speed of light is 120ms, so if the \ac{UE} sends a request to a server at time zero through a \ac{GEO} link, the best-case delay will be of 480ms, taking into account only the propagation delay.
    The large cell footprint also means that a single satellite will be serving a massive number of users, so the total available capacity will have to be shared between more users, threfore the throughput experienced by each one of them will be reduced.
\end{itemize}

\subsection{Current solutions}
Current solutions for non-terrestrial communication do exist, but they mostly rely on telecommunications satellites placed in the \ac{GEO}, at a height of 35.786 km. The distance that the signal has to travel offering limited throughput and large delays. While \ac{LEO} constellations (400 km to 2.000 km) have proven to be a valid alternative, providing higher throughput and lower latency \cite{main-features-5g-nr-ntn-yun}, they have the drawback of an increased Doppler shift due to their high speed relative to ground \cite{satellite-communication-mmwave-giordani}, and there is still no international standard with regard to the communication protocols to use. 

\paragraph{}
This scenario led \ac{3GPP} to identify some work to be done to integrate \ac{NTN} in cellular standards, calling for long-term research in this field \cite{satellite-communication-mmwave-giordani}. This work will mainly focus on the 

\paragraph{}\todo{move this to state of the art} Focusing on the \ac{MAC} sublayer, the large propagation delay of satellite links affects different aspects, making the actual implementation not suited for a \ac{NTN} scenario. In the \ac{HARQ} protocol, the retransmission timeout is likely to expire before a single \ac{RTT}, leading to unnecessary retransmissions. Moreover, the limit on the maximum number of concurrent \ac{HARQ} processes leads to a stop-and-wait behaviour, which may increase the energy consumption \cite{3gpp-tr-38.811}. On the other hand, it has been noted that disabling \ac{HARQ} would lead to an even worse performance penalty, therefore requiring a redesign for \ac{NTN} \cite{5g-beyond-5g-ntn-trends-vanellicoralli}. Another 5G \ac{NR} protocol which is negatively impacted in \ac{NTN} is the initial access, since users at the centre of the cell face a smaller propagation delay with respect to users at the cell edge \cite{5g-beyond-5g-ntn-trends-vanellicoralli} \cite{applying-nr-technologies-in-ntn-lee}. As a result, preambles of \ac{UE}s placed near the cell edge may reach the satellite when the \ac{RACH} opportunity has already expired, which may lead to collisions. During the initial access phase, \ac{UE}s are not aware of their propagation delay, and the high mobility of \ac{gNB}s on \ac{LEO} satellites causes a non-negligible Doppler shift. Those factors vary with the relative position and speed between the \ac{UE} and the \ac{gNB}, and the protocols for initial access must be modified in \ac{NTN} to account for them \cite{ntn-from-5g-6g-hassan}. 
\paragraph{}
It is clear that the future of mobile networks envisioned by \ac{3GPP} embraces \ac{NTN}s, and considerable work has to be done. Research will bear a high impact towards a more connected, equal opportunity world. 

\section{Currrent state of the art}
\todo{types of payloads: regenerative, and transparent or bent pipe}